---
title: "Weather"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

## Introduction  
   
In this analysis, we investigate the relationship between the daily minimum temperature (MIN) and the daily maximum temperature (MAX) for two weather stations: one in Rexburg and one in Salt Lake City. The motivation is to understand how well the daily (MIN) temperature predicts the daily (MAX) temperature and to see if the relationship differs between the two locations. To accomplish this, we use the **GSODR** package to retrieve weather data for the year 2021, perform a multiple linear regression analysis, and visually inspect the model with various plots.  
   
## Setup and Data Acquisition 

```{r, warning=FALSE, message=FALSE, ev}
# Load necessary packages  
library(GSODR)  # Ensure this package is installed (install.packages("GSODR") if needed)  
library(dplyr)  
library(ggplot2)  
library(car)  
library(DT)  
library(tidyr)  
library(mosaic)
library(pander)
  
# Load additional data if needed (isd_history might help to identify stations)  
data("isd_history")  
  
# Choose two weather stations  
# For example: Rexburg -- STNID = "726818-94194" and Salt Lake City -- STNID = "725720-24127"  
  
# Download data for 2021  
Rexburg <- get_GSOD(years = 2021, station = "726818-94194")  
Salt_Lake <- get_GSOD(years = 2021, station = "725720-24127")  
  
# Review downloaded data if desired  
# View(Rexburg)  
# View(Salt_Lake)  
  
# Combine the two datasets  
weather <- rbind(Rexburg, Salt_Lake)  
  
# Select only the relevant columns for the analysis: station NAME, MAX, and MIN temperatures  
weather2 <- weather %>% select(NAME, MAX, MIN)  
  
# Optional: show table preview  
datatable(weather2, options = list(pageLength = 5)) 
```

## Hypotheses and Significance Level
We test whether the daily minimum temperature and the weather station location significantly predict the daily maximum temperature.

__Null Hypothesis (H₀)__: The minimum temperature and station location have no effect on the maximum temperature (i.e., all regression coefficients = 0).

__Alternative Hypothesis (H₁)__: At least one regression coefficient (MIN or NAME) is not equal to 0, indicating a significant relationship.

__Significance Level (α):__ 0.05

## Models:

The fitted multiple linear regression model predicts the daily maximum temperature (MAX) using the daily minimum temperature (MIN) and the station location (Salt Lake City or Rexburg). The model is expressed as:

$$
\text{MAX}_i = 15.63 + 1.235 \cdot \text{MIN}_i - 3.143 \cdot \text{SaltLake}_i + \varepsilon_i
$$

In this equation, Rexburg serves as the reference category for the station variable. The intercept (15.63) represents the expected maximum temperature in Rexburg when the minimum temperature is 0°C. The coefficient for MIN (1.235) indicates that for each 1°C increase in minimum temperature, the maximum temperature is expected to increase by approximately 1.235°C, assuming the location remains constant. The coefficient for Salt Lake City (-3.143) suggests that, on average, Salt Lake City's maximum temperatures are about 3.14°C lower than Rexburg’s when both locations have the same minimum temperature. The error term accounts for random variation not explained by the model.

## Graphical Summary

We now create a scatter plot of MIN versus MAX colored by station name. A linear regression line is added for each group to visually assess the relationship. This plot gives an initial overview of the linear association between the minimum and maximum temperatures for each city.

```{r, message=FALSE}
# Scatter plot with separate regression lines for each station  
ggplot(weather2, aes(x = MIN, y = MAX, color = NAME)) +  
 geom_point(alpha = 0.6) +  
 geom_smooth(method = 'lm', se = FALSE) +  
 labs(title = 'Scatter Plot of Maximum vs Minimum Temperature',  
      x = 'Minimum Temperature (°C)',  
      y = 'Maximum Temperature (°C)') +  
 theme_minimal()  
```

## Numerical Summary: Multiple Linear Regression

Next, we fit a multiple linear regression model predicting MAX as a function of MIN and the station (represented by NAME). This allows us to evaluate both the overall effect of MIN and potential differences between the two locations.

```{r}
# Fit the multiple linear regression model  
weather_lm <- lm(MAX ~ MIN + NAME, data = weather2)  
 
# Summary of the regression model  
pander(summary(weather_lm)) 
```

The output of the regression summary will indicate the estimated coefficients, standard errors, t-statistics, and p-values for each predictor. This helps us understand the significance and magnitude of the relationship between MIN and MAX across the two cities.

## Diagnostic Plots for Model Requirements

To evaluate if the assumptions of the linear regression model are satisfied, we generate several diagnostic plots. These include:

__Residuals vs Fitted Plot:__ Checks for issues with non-linearity and unequal variance.
__Normal Q-Q Plot:__ Assesses normality of the residuals.
__Scale-Location Plot:__ Verifies homoscedasticity.
__Residuals vs Leverage Plot:__ Identifies influential observations.

```{r}
# Diagnostic plots  
par(mfrow = c(2,2))  
plot(weather_lm)  
```

Additionally, an added variable plot is used to show the contribution of the MIN variable after accounting for the effect of NAME.

```{r}
# Added variable plot for MIN  
avPlot(weather_lm, variable = 'MIN', main = 'Added Variable Plot for MIN')  
```

## Conclusion and Interpretation
Based on the regression results, we can draw several important conclusions about the relationship between minimum and maximum temperatures across the two weather stations.

The multiple linear regression model (MAX ~ MIN + NAME) demonstrates a strong relationship between the variables, with an adjusted R-squared value of 0.8613. This indicates that approximately 86.13% of the variation in maximum temperature can be explained by the minimum temperature and the station location, which represents an excellent fit for weather data.

Looking at the specific coefficients:

- The intercept (15.63°C) represents the expected maximum temperature in Rexburg when the minimum temperature is 0°C. This baseline is statistically significant (p < 6.01e-308).

- The coefficient for MIN (1.235) indicates that for each 1°C increase in minimum temperature, we expect the maximum temperature to increase by approximately 1.235°C, holding the location constant. This relationship is highly significant (p < 3.227e-308).

- The coefficient for NAME (SALT LAKE CITY INTERNATIONAL AIRPORT) is -3.143, which means that Salt Lake City's maximum temperatures are, on average, 3.143°C lower than Rexburg's when controlling for the same minimum temperature. This difference is statistically significant (p < 1.035e-17).

- The residual standard error of 4.544°C represents the average deviation of observed maximum temperatures from the predicted values. While this indicates some variability not captured by our model, the high R-squared value suggests that the model is still quite effective at predicting maximum temperatures.

In practical terms, these results suggest that Rexburg experiences more extreme temperature swings from minimum to maximum compared to Salt Lake City. This could be attributed to various geographical factors such as elevation, proximity to water bodies, or urban heat island effects. The strong predictive power of minimum temperature across both locations highlights its utility as a forecasting tool for daily maximum temperatures.



