---
title: "Logistic-Analysis"
output: html_document
---

```{r}
pacman::p_load(mosaic, pander, tidyverse, car, lattice, DT, alr4, ResourceSelection)
```
1) Background/Context
In this analysis, we investigate how the property area (in m²) predicts whether a property is expensive (price above the median). The dataset comprises various features like property type, state, area, and price in USD.

```{r}
# Load the dataset  
real_estate <- read.csv('mexico-real-estate-clean.csv', stringsAsFactors = TRUE)  
  
# Display the first few rows of the dataset  
print(head(real_estate))  
  
# Create a binary outcome: 'expensive' is 1 if price is above the median, else 0  
median_price <- median(real_estate$price_usd, na.rm = TRUE)  
real_estate$expensive <- ifelse(real_estate$price_usd > median_price, 1, 0)  
real_estate$expensive <- as.factor(real_estate$expensive)  
```

2) Hypothesis Testing with Model
Null Hypothesis (H₀): The coefficient for area_m2 equals 0 (i.e., no effect on the probability of being expensive).
Alternative Hypothesis (H₁): The coefficient for area_m2 is nonzero (i.e., area influences the probability).

```{r}
# Fit logistic regression model using area_m2 as the predictor  
model <- glm(expensive ~ area_m2, data = real_estate, family = binomial)  
summary_model <- summary(model)  
```

3) Graph
The following plot shows the observed data and the fitted logistic regression curve.

```{r}
# Create a sequence for area_m2 and predict probabilities  
area_seq <- seq(min(real_estate$area_m2, na.rm = TRUE), max(real_estate$area_m2, na.rm = TRUE), length.out = 100)  
pred_frame <- data.frame(area_m2 = area_seq)  
pred_frame$pred_prob <- predict(model, newdata = pred_frame, type = 'response')  
  
# Plot using ggplot2  
plot_lr <- ggplot(real_estate, aes(x = area_m2, y = as.numeric(as.character(expensive)))) +   
  geom_jitter(height = 0.05, alpha = 0.5) +   
  geom_line(data = pred_frame, aes(x = area_m2, y = pred_prob), color = 'blue', size = 1) +   
  labs(title = 'Logistic Regression: Probability of Expensive Property vs Area (m²)',  
       x = 'Area (m²)',  
       y = 'Probability of Being Expensive')  
print(plot_lr)  
```

4) Analysis
The following output displays the logistic regression model summary, including the estimated coefficients, standard errors, and p-values.
```{r}
# Display model summary using pander for formatting  
pander(summary_model)  
```
5) Prediction
We now predict the probability of a property with an area of 150 m² being expensive.
```{r}
# Predict probability for a new property with area of 150 m²  
new_property <- data.frame(area_m2 = 150)  
predicted_prob <- predict(model, newdata = new_property, type = 'response')  
cat('Predicted probability for a property with 150 m² is:', predicted_prob, '\n')  
```
6) Explanation of exp
In logistic regression, exp(b1) represents the odds ratio for a one-unit increase in the predictor variable. Here, it indicates the multiplicative change in the odds of a property being expensive when the property area increases by 1 m².
```{r}
cat("exp(b1), the odds ratio for area_m2, indicates the multiplicative change in the odds of a property being expensive for each additional m².\n")
```
7) Goodness of Fit Test
We perform the Hosmer-Lemeshow goodness of fit test using the ResourceSelection package.

```{r}
hoslem_result <- hoslem.test(as.numeric(as.character(real_estate$expensive)), fitted(model), g = 10)  
pander(hoslem_result)  
```

8) Conclusion/Future Studies
The logistic regression analysis indicates that property area is a significant predictor of whether a property is expensive. Specifically, as the area increases, the probability of the property being expensive increases.

Future studies could include:

Additional predictors such as location, property type, and neighborhood characteristics.
Investigation of interaction effects between predictors.
Exploration of non-linear relationships.
Analysis across different segments or regions.

```{r}
cat("Conclusion: The logistic regression model suggests that property area (m²) is a significant predictor of a property being expensive.\n")  
cat("Future studies could incorporate additional predictors like location and property type to further improve model performance.\n")  
```



