---
title: "Rent"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
    toc: true
    toc_float: true
---

<!-- Welcome to the first Analysis for Math 325! 
The goal of this first assignment is to have you produce a graphical summary and a numerical summary of the "Rent" data set. The fun part is that you get to decide what type of graph and what type of numerical summary to create! You will need to read through the "green text" instructions of this document to find the places where you are supposed to enter R-code in order to complete the assignment. -->


**Week 1 Analysis**

```{r, message=FALSE, warning=FALSE}
# This gray area is called an "R-chunk". 
# These library commands install some powerful functions for your use later on.
library(mosaic)
library(pander)
library(tidyverse)
library(DT) 

# This read_csv command reads in the "Rent" data set into an object called "Rent"
Rent <- read_csv("../Data/Rent.csv")

# To load this data set into your R-Console do the following:
#  1. From your top file menu select "Session -> Set Working Directory -> To Source File Location"
#  2. Press the green "play button" in the top right corner of this gray box (which is called an "R-chunk").
#  3. Then in your "Console" window of 
```

<br />

## Background

Here is a data table showing the available approved housing apartment options at BYU-Idaho for single students. There are 110 entries comprising 54 female and 56 male apartment options.

```{r, warning=FALSE}
# Code to get you started. 
# View(...) works great in the Console, but datatable(...) must be
# used instead within an R-chunk.
datatable(Rent, options=list(lengthMenu = c(3,10,30)), extensions="Responsive")
```


## Graphic

<!-- Present an interesting graphic using the Rent data set. Do not create the same graphics that are shown in the [Good Example Analysis](./Analyses/StudentHousing.html), but you can certainly use those graphics as a guide to inspire you on what to do yourself. -->


```{r}
# Use this R-chunk to...
# ...Create an interesting graphic using the Rent data set.

library(ggplot2)

ggplot(Rent, aes(x = CrowFlyMetersToRicks, y = AvgFloorPlanCost, color = AvgFloorPlanCost)) +
         geom_point()
```


<!-- Write a few statements here that explain your graphic and what it shows. -->

This graph shows the relationship between distance of apartment to Ricks building and the average floor plan cost. You might think that the closer you are to campus, the more expensive it gets. However, that's not always the case. In fact, there is no relationship at all. You can live close to campus and pay an expensive or cheap rent. 


```{r}
# Use this R-chunk to...
# ...compute and display a meaningful table of numerical summaries supporting your above graphic.

summary(Rent$AvgFloorPlanCost)
```

<!-- Write a few statements here that introduce your table of numerical summaries and what they show. -->

The summary statistics shown above shows the average floor plan costs. The average floor plan cost is $1324, with minimum rent of $880 and highest of $1790.


## Conclusion

If you are a data science student, you might want to live close to Ricks Building, especially during Winter semester when walking to campus is difficult. However, don't just take the first close apartment you see. You might be paying premium and go broke at an instant. Well, you can do that if you are rich, of course. However, if you are like me, who try to make ends meet every semester, try to look around for better prices. You can pay as low as $880 and save the $500 every semester. If you study for 8 semesters, that's instant $4000 savings for you.



