---
title: "Food at College"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

```{r, warning=FALSE, message=FALSE}
# From your file menu (at the top of R-Studio) select:
# "Session -> Set working directory -> To source file location"
# Then play this chunk to get the data into R.
library(mosaic)
library(car)
library(DT)
library(pander)
library(readr)
food <- read_csv("~/Statistics-Notebook-Github/Data/food.csv") #food.csv is in the Data folder...
```

## 1. Context

This analysis investigates the relationship between groups defined by comfort food reasons (coded) and daily calorie intake. The analysis uses data from the `food.csv` file. Specifically, we are interested in whether different groups (as indicated by the variable `comfort_food_reasons_coded`) have significantly different average calorie intakes (`calories_day`).

## 2. Hypotheses for One-way ANOVA

- $$H_0: \mu_1 = \mu_2 = \dots = \mu_k$$ 
- $$H_A: \	ext{At least one group mean is different}$$

where $$\mu_i$$ represents the mean daily calorie intake for group $$i$$.

## 3. Numerical and Graphical Summaries

### Numerical Summary

```{r numeric-summary}
# Clean data: Remove rows with missing values in key columns
food_clean <- food %>% 
  filter(!is.na(calories_day) & !is.na(comfort_food_reasons_coded))

# Generate summary statistics
summary_stats <- food_clean %>% 
  group_by(comfort_food_reasons_coded) %>% 
  summarise(
    count = n(),
    mean_cal = mean(calories_day, na.rm = TRUE),
    sd_cal = sd(calories_day, na.rm = TRUE)
  )
summary_stats
```

### Graphical Summary

```{r graphic-summary}
# Boxplot of calories_day by comfort_food_reasons_coded
p <- ggplot(food_clean, aes(x = factor(comfort_food_reasons_coded), y = calories_day)) + 
  geom_boxplot(fill = "lightblue") + 
  labs(x = "Comfort Food Reasons (Coded)", y = "Calories per Day", 
       title = "Calories per Day by Comfort Food Reason Groups") + 
  theme_minimal()
print(p)
```

## 4. One-way ANOVA

```{r anova}
# Fit one-way ANOVA model
anova_model <- aov(calories_day ~ as.factor(comfort_food_reasons_coded), data = food_clean)
summary(anova_model)
```

## 5. Check Requirements

Before relying on ANOVA, we need to check its assumptions:
- **Normality:** Residuals should be normally distributed.
- **Homogeneity of Variance:** Different groups should have similar variances.

```{r assumptions-check}
# Extract residuals
residuals_anova <- residuals(anova_model)

# Normality check (Shapiro-Wilk test)
shapiro_test <- shapiro.test(residuals_anova)
print(shapiro_test)

# Homogeneity of Variance (Levene\'s Test from the car package)
library(car)
levene_test <- leveneTest(calories_day ~ as.factor(comfort_food_reasons_coded), data = food_clean)
print(levene_test)
```

## 6. Testing Approach

If the ANOVA assumptions are met, we will use the ANOVA results above. If not, we will use the non-parametric Kruskal-Wallis test as follows.

```{r kruskal-wallis}
kw_test <- kruskal.test(calories_day ~ as.factor(comfort_food_reasons_coded), data = food_clean)
print(kw_test)
```

## 7. Conclusion and Future Studies

Based on the tests, we conclude the following:

- If the ANOVA assumptions are met, the one-way ANOVA results can be used to infer if there are significant differences in calorie intake between the groups.

- If the assumptions are not met, the Kruskal-Wallis test provides a valid non-parametric alternative.

Future studies might consider further data cleaning, including additional predictors, or collecting more data to increase the study\'s power.

*End of Analysis.*'

# Save the RMD document
writeLines(rmd_text, con = 'analysis.Rmd')

# Display the first few lines of the RMD file
file_conn <- file('analysis.Rmd', open = 'r')
rmd_preview <- readLines(file_conn, n = 20)
close(file_conn)
print(rmd_preview)






