---
title: "Food at College"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---
  
```{r, warning=FALSE, message=FALSE}
  # From your file menu (at the top of R-Studio) select:
  # "Session -> Set working directory -> To source file location"
  # Then play this chunk to get the data into R.
library(mosaic)
library(car)
library(DT)
library(pander)
library(readr)
food <- read_csv("food.csv") #food.csv is in the Data folder...
```
  
  <!-- Instructions:
  
   Use the Food dataset and a Kruskal-Wallis Test(s) to answer an interesting question(s) that you come up with.
   
   About the dataset:
   
    * GPA - numerical, self-reported current college GPA
  
    
    * weight - numerical, self-reported weight in lbs
  
  
    * gender - "Female", "Male"
  
  
    * breakfast - "Cereal", "Donut" 
        The participants were shown a picture of "Cereal" and a "Donut" and asked which one of these pictures they associate with the word "breakfast."
  
  
    * calories_chicken - guessing calories in chicken piadina from the options of:
       1 - 265 
       2 - 430 
       3 - 610 
       4 - 720
      (the variable shows the actual number of calories participants selected)
  
  
     * calories_day - Importance of consuming calories per day
       1 - i dont know how many calories i should consume 
       2 - it is not at all important 
       3 - it is moderately important 
       4 - it is very important
  
  
     * comfort_food - List 3-5 comfort foods that come to mind. 
       Open ended 
  
  
     * comfort_food_reasons - What are some of the reasons that make you eat comfort food?      (i.e., anger, sadness, happiness, boredom, etc) - list up to three
       Open ended
  
  
     * comfort_food_reasons_coded
       1 - stress
       2 - boredom
       3 - depression/sadness
       4 - hunger
       5 - laziness
       6 - cold weather
       7 - happiness 
       8 - watching tv
       9 - none 
  
  
      * cook - how often do you cook?
       1 - Every day 
       2 - A couple of times a week 
       3 - Whenever I can, but that is not very often  
       4 - I only help a little during holidays 
       5 - Never, I really do not know my way around a kitchen
  
  
      * cuisine - what type of cuisine did you eat growing up?
       1 - American
       2 - Mexican.Spanish
       3 - Korean/Asian
       4 - Indian
       5 - American inspired international dishes
       6 - other
  
  
      * diet_current - describe your current diet
       open ended
  
  	
      * diet_current_coded
       1 - healthy/balanced/moderated/
       2 - unhealthy/cheap/too much/random/
       3 - the same thing over and over
       4 - unclear
  
  
      * drink - which picture do you associate with the word “drink”?
       1 - orange juice
       2 - soda 
  
  
      * eating_changes  - Describe your eating changes since the moment you got into college?
       Open ended 
  
  
      * eating_changes_coded
       1 - worse
       2 - better
       3 - the same
       4 - unclear
  
  
      * eating_changes_coded1
       1 - eat faster
       2 - bigger quantity
       3 - worse quality 
       4 - same food
       5 - healthier
       6 - unclear
       7 - drink coffee 
       8 - less food
       9 - more sweets
       10 - timing 
       11 - more carbs or snacking
       12 - drink more water
       13 - more variety
  
  
      * eating_out - frequency of eating out in a typical week 
       1 - Never 
       2 - 1-2 times 
       3 - 2-3 times 
       4 - 3-5 times 
       5 - every day
  
  
      * employment - do you work? 
       1 - yes full time 
       2 - yes part time 
       3 - no
       4  - other
  
  
      * ethnic_food - How likely to eat ethnic food 
       1 - very unlikely 
       2 - unlikely 
       3 - neutral 
       4 - likely 
       5 - very likely 
  
  
      * exercise - how often do you exercise in a regular week?
       1 - Everyday 
       2 - Twice or three times per week 
       3 - Once a week
       4 - Sometimes 
       5 - Never
  
  
      * father_education - 
       1 - less than high school 
       2 - high school degree 
       3 - some college degree 
       4 - college degree 
       5 - graduate degree 
  
  
      * father_profession - what is your father profession?
       Open ended
  
  
      * fav_cuisine - What is your favorite cuisine?
       Open ended
  
  
      * fav_cuisine_coded
       0-none
       1 - Italian/French/greek
       2 - Spanish/mexican
       3 - Arabic/Turkish
       4 - asian/chineses/thai/nepal
       5 - American
       6 - African 
       7 - Jamaican
       8 - indian
  
  
      * fav_food - was your favorite food cooked at home or store bought? 
       1 - cooked at home 
       2 - store bought 
       3 - both bought at store and cooked at home
  
  
      * food_childhood - what was your favorite childhood food?
       Open ended
  
  
  
      * fries - which of these pictures you associate with word fries? 
       1 - Mcdonald’s fries
       2 - home fries
  
   
      * fruit_day - How likely to eat fruit in a regular day 
       1 - very unlikely 
       2 - unlikely 
       3 - neutral 
       4 - likely 
       5 - very likely 
  
  
      * grade_level – 
       1 - freshman 
       2 -Sophomore 
       3 - Junior 
       4 - Senior
       
        
      * greek_food - How likely to eat greek food when available?
       1 - very unlikely 
       2 - unlikely 
       3 - neutral 
       4 - likely 
       5 - very likely 
       
        
      * healthy_feel – how likely are you to agree with the following statement: “I feel very healthy!” ?
        1 to 10 where 1 is strongly agree and 10 is strongly disagree - scale
        
        
      * healthy_meal – what is a healthy meal? Describe in 2-3 sentences.
        Open ended
        
        
      * ideal_diet – describe your ideal diet in 2-3 sentences
        Open ended
        
        
      * Ideal_diet_coded
        1 – portion control
        2 – adding veggies/eating healthier food/adding fruit	
        3 – balance
        4 – less sugar
        5 – home cooked/organic
        6 – current diet
        7 – more protein
        8 – unclear
        
        
       * income
        1 - less than $15,000 
        2 - $15,001 to $30,000 
        3 - $30,001 to $50,000 
        4 - $50,001 to $70,000 
        5 - $70,001 to $100,000 
        6 - higher than $100,000
        
       * indian_food – how likely are you to eat indian food when available
        1 - very unlikely 
        2 - unlikely 
        3 - neutral 
        4 - likely 
        5 - very likely 
        
        
       * Italian_food – how likely are you to eat Italian food when available?
        1 - very unlikely 
        2 - unlikely 
        3 - neutral 
        4 - likely 
        5 - very likely 
        
        
       * life_rewarding – how likely are you to agree with the following statement: “I feel life is very rewarding!” ?
        1 to 10 where 1 is strongly agree and 10 is strongly disagree - scale
        
        
       * marital_status
        1 -Single 
        2 - In a relationship 
        3 - Cohabiting 
        4 - Married 
        5 - Divorced 
        6 - Widowed
        
        
       * meals_dinner_friend – What would you serve to a friend for dinner?
        Open ended
        
        
       * mothers_education 
        1 - less than high school 
        2 - high school degree 
        3 - some college degree 
        4 - college degree 
        5 - graduate degree
        
        
       * mothers_profession – what is your mother’s profession? 
        
        
       * nutritional_check - checking nutritional values frequency 
        1 - never 
        2 - on certain products only 
        3 - very rarely 
        4 - on most products 
        5 - on everything
        
        
       * on_off_campus – living situation
        1 - On campus 
        2 - Rent out of campus 
        3 - Live with my parents and commute 
        4 - Own my own house
        
        
       * parents_cook - Approximately how many days a week did your parents cook? 
        1 - Almost everyday 
        2 - 2-3 times a week 
        3 - 1-2 times a week 
        4 - on holidays only 
        5 - never
        
        
       * pay_meal_out - How much would you pay for meal out? 
        1 - up to $5.00 
        2 - $5.01 to $10.00 
        3 - $10.01 to $20.00 
        4 - $20.01 to $30.00 
        5 - $30.01 to $40.00 
        6 - more than $40.01 
        
        
       * Persian_food - How likely to eat persian food when available?
        1 - very unlikely 
        2 - unlikely 
        3 - neutral 
        4 - likely 
        5 - very likely 
        
        
       * self_perception_weight - self perception of weight 
        6 - i dont think myself in these terms 
        5 - overweight 
        4 - slightly overweight 
        3 - just right 
        2 - very fit 
        1 - slim 
        
        
       * Which of the two pictures you associate with the word soup?
        1 – veggie soup
        2 – creamy soup
        
        
       * sports - sports – do you do any sporting activity?
        1 - Yes 
        2 - No 
        99 – no answer
        
        
       * thai_food - How likely to eat thai food when available?
        1 - very unlikely 
        2 - unlikely 
        3 - neutral 
        4 - likely 
        5 - very likely 
        
        
       * tortilla_calories - guessing calories in a burrito sandwhich from Chipolte?
        1 - 580 
        2 - 725 
        3 - 940 
        4 - 1165
        
       * turkey_calories - Can you guess how many calories are in the foods shown below?
       (Panera Bread Roasted Turkey and Avocado BLT)
        1 - 345 
        2 - 500 
        3 - 690 
        4 - 850
        
        
       * type_sports – what type of sports are you involved?
        Open-ended
        
       * veggies_day - How likely to eat veggies in a day? 
        1 - very unlikely 
        2 - unlikely 
        3 - neutral 
        4- likely 
        5 - very likely
        
       * vitamins – do you take any supplements or vitamins?
        1 – yes
        2 – no
        
       * waffle_calories - guessing calories in waffle potato sandwhich 
        1 - 575 
        2 - 760 
        3 - 900 
        4 - 1315
  
  
  Note this dataset and description come from: © 2020 Kaggle Inc for further details visit:
  
  [Food choices on Kaggle](https://www.kaggle.com/borapajo/food-choices?select=food_coded.csv)
  
  
  If you want to give your critiquers some ideas about what you have questions on and would like help with, place those details here.-->
  
  
  
  
  <!--End your comments to your critiquers before this line.-->
  
  ----
  
  <!--Begin writing your Analysis after this point.-->
  
## Background
## 1. Context
This analysis investigates the relationship between how healthy individuals feel (measured by the healthy_feeling variable) and how often their parents cooked during their childhood (measured by parents_cook). Understanding this relationship could shed light on potential long-term influences of early home-cooked meals on health perceptions later in life.
  
## 2. Hypotheses: One-Way ANOVA
We set up a one-way ANOVA to test whether the mean perceived healthy feeling differs by parental cooking frequency.
  
\\begin{aligned}H_0&: \\mu_1 = \\mu_2 = \\dots = \\mu_k\\\\H_A&: \\text{At least one group mean is different.}\\end{aligned}
  
## 3. Numerical and Graphical Summaries
  
```{r}
  # Load data
  df <- read.csv("food.csv")
  
  df <- df %>%
    mutate(healthy_feeling = case_when(
      healthy_feeling == 1 ~ 10,
      healthy_feeling == 2 ~ 9,
      healthy_feeling == 3 ~ 8,
      healthy_feeling == 4 ~ 7,
      healthy_feeling == 5 ~ 6,
      healthy_feeling == 6 ~ 5,
      healthy_feeling == 7 ~ 4,
      healthy_feeling == 8 ~ 3,
      healthy_feeling == 9 ~ 2,
      healthy_feeling == 10 ~ 1,
      TRUE ~ healthy_feeling  # catch any other values
    )) %>%
    filter(
      parents_cook < 4
    )
  
# Summarize healthy_feeling by parents_cook with reversed scoring
summary_table <- df %>%
  group_by(parents_cook) %>%
  summarise(Count = n(), 
              Mean = round(mean(healthy_feeling, na.rm = TRUE), 2), 
              SD = round(sd(healthy_feeling, na.rm = TRUE), 2))

pander(summary_table)
  
# Create boxplot using ggplot2
p <- ggplot(df, aes(x = factor(parents_cook), y = healthy_feeling)) +
  geom_boxplot(fill = 'lightblue', color = 'darkblue') +
  labs(title = "Healthy Feeling by Parental Cooking Frequency",
        x = "Parents Cook (Frequency)",
        y = "Healthy Feeling") +
  theme_minimal()

print(p)
```

Based on the numerical and graphical summaries, it appears individuals whose parents cooked once or twice a week during their childhood feel healthier in adulthood.
  
## Anova Test
```{r}
# Perform one-way ANOVA
  
anova_mod <- aov(healthy_feeling ~ factor(parents_cook), data = df)
anova_summary <- summary(anova_mod)
pander(anova_summary)
```

The ANOVA test results tell us that even though we saw some differences in the average "healthy_feel" scores across the different parental cooking frequency groups, those differences are likely due to random chance. There's no strong statistical evidence to say that how often your parents cooked when you were a kid has a significant impact on how healthy you feel now, based on this data.
  
## Residual Plot
```{r}
# Residual plot for ANOVA model
model_resid <- residuals(anova_mod)
plot(df$parents_cook, model_resid,
      main = "Residuals vs Parents Cook Frequency",
      xlab = "Parents Cook (Frequency)",
      ylab = "Residuals")
abline(h = 0, col = "red")
```
  
This residual plot, displaying residuals against parental cooking frequency, is used to assess the assumptions of the ANOVA. Ideally, residuals should be randomly scattered around zero, indicating that the model fits the data well and that the variance is consistent across groups. In this plot, while there's a general scattering of points, there's a noticeable pattern: residuals tend to cluster around specific vertical lines corresponding to each parental cooking frequency category (1, 2, and 3), suggesting potential issues with the assumption of homogeneity of variances. Additionally, there are a few outliers, especially at the lower end of the residual scale, which could further impact the validity of the ANOVA results. This pattern indicates that the model might not be capturing all the variability in the data, and further analysis or alternative statistical methods might be warranted.
  
## QQPlot
```{r}
qqnorm(model_resid, main = "QQ Plot of ANOVA Residuals")
qqline(model_resid, col = "red")
```

This QQ (Quantile-Quantile) plot of ANOVA residuals is used to assess the normality assumption of the residuals. Ideally, if the residuals are normally distributed, the points on the plot should fall closely along the red diagonal line. In this plot, while the central points generally follow the line, there are noticeable deviations at both ends. Specifically, the points at the lower and upper tails of the distribution deviate from the straight line, suggesting that the residuals might not be perfectly normally distributed. This indicates potential violations of the normality assumption, which could affect the reliability of the p-values from the ANOVA test. While moderate deviations might not invalidate the results entirely, significant deviations, as observed here, suggest that caution should be exercised when interpreting the ANOVA results.
  
## 6. Comparison with Non-Parametric Test
Based on the residual and QQ plots, since the requirements are not met, we perform a Kruskal-Wallis test as a non-parametric alternative.
  
## 7. Kruskal-Wallis Test
__Hypotheses__

__Null hypothesis (H₀)__: The distribution of healthy_feeling scores is the same across all parental cooking frequency groups.

__Alternative hypothesis (H₁)__: At least one group differs in the distribution of healthy_feeling scores.

Since the assumptions for ANOVA were not met (as seen in the residual and QQ plots), the Kruskal-Wallis test provides a more appropriate, non-parametric method for comparing multiple groups when normality and homogeneity of variances are questionable.
  
```{r}
kruskal_result <- kruskal.test(healthy_feeling ~ factor(parents_cook), data = df)
pander(kruskal_result)
```
  
The Kruskal-Wallis test produced a test statistic of **0.9285** with **2 degrees of freedom**, resulting in a **p-value of 0.6286**. Since this p-value is greater than the significance level of 0.05, we **fail to reject the null hypothesis**. This means we do not have sufficient evidence to conclude that perceived healthiness differs based on how often parents cooked during the respondent's childhood.
  
## 8. Conclusion and Future Studies

This analysis aimed to explore whether the frequency of parental cooking during childhood influences how healthy college students feel as adults.

While descriptive statistics suggested that those whose parents cooked 1–2 times a week reported slightly higher average health ratings (Mean ≈ `r summary_table$Mean[summary_table$parents_cook == 3]`), this trend was not statistically significant.

- **ANOVA Result:**  
  F(2, 121) = 0.462, p = **0.6313**
  
- **Kruskal-Wallis Result:**  
  Test statistic = 0.9285, p = **0.6286**

Because both tests yielded p-values well above 0.05, we **fail to reject the null hypothesis** in each case. This suggests that **any observed differences in perceived healthiness across groups are likely due to random variation** rather than a meaningful association with parental cooking frequency.

### Suggestions for Future Research:
- Include a broader range of cooking frequency options (e.g., actual number of meals per week).
- Investigate mediating variables such as current eating habits, meal quality, or family health culture.
- Collect a larger sample to improve the power of statistical tests.

Despite not finding statistical significance, the process highlighted interesting patterns and raised new questions for further exploration in understanding the long-term effects of childhood eating environments.
