---
title: "All Class Activities"
format: html
editor: visual
---

```{r}
pacman::p_load(mosaic, pander, tidyverse, car, lattice, DT, alr4, ResourceSelection)
```

## Week 2 Day 1

### Graphical Summaries

```{r}
# Libraries used: MOSAIC, PANDER, TIDYVERSE
# Dataset:        KIDSFEET

# Group 1
boxplot(length~sex,data=KidsFeet)

KidsFeet %>% 
  group_by(sex) %>% 
  summarise(min = min(length), 
            Q1=quantile(length,0.25),
            median=median(length),
            Q3=quantile(length,0.75),
            max=max(length))

favstats(length~sex,data=KidsFeet)

# Group 2
barplot(table(KidsFeet$sex))

table(KidsFeet$sex)

# Group 3
plot(length~width, data=KidsFeet)
       
KidsFeet %>% 
  summarise(Correlation=cor(width,length))
```

## Week 2 Day 2

### Graphical Summaries

```{r}
# Libraries used: MOSAIC, TIDYVERSE
# Dataset:        AIRQUALITY

# Histograms (length)
hist(airquality$Wind,main="La Guardia Airport from May to September, 1973", 
     xlab="Daily Average Wind Speeds (mph)",
     ylab="Frequency",
     col="steelblue",
     breaks = 22)

ggplot(airquality, aes(x=Wind))+
  geom_histogram(fill="steelblue")+
  labs(title="La Guardia Airport from May to September, 1973",
       x="Daily Average Wind Speeds (mph)",
       y="Frequency")

# boxplots
boxplot(Wind~Month,main="Wind by Month", xlab="Month",
     ylab="Wind",col=c("steelblue1", "steelblue2", "steelblue3","steelblue3",
                       "steelblue2"),data=airquality)

ggplot(data=airquality,aes(y=Wind,x=Month,group=Month))+
  geom_boxplot(fill=c("steelblue1", "steelblue2", "steelblue3","steelblue3",
                      "steelblue2"))+
  labs(title = "Wind by Month",
       y="Wind",
       x="Month")

# scatter plots 
plot(Ozone~Temp,data=airquality,
     main="Relationship Between Ozone and Temperature",
     xlab="Temp", 
     ylab="Ozone",
     pch = 19, 
     col="blue")

ggplot(airquality,aes(x=Temp,y=Ozone))+
  geom_point(color="blue")+
  labs(title="Relationship Between Ozone and Temperature", 
       x="Temp", 
       y="Ozone")

## Go over the Stephanie analysis
```

## Week 3 Day 1

### Graphical and Numerical Summaries

```{r}
# Libraries used: MOSAIC, PANDER, TIDYVERSE
# Dataset:        KIDSFEET

#Day 1
# A) Quantitative vs Categorical Data - KidsFeet
# B) Statistics-Notebook page - Identify columns where a dataset would match "data logos"

# One Quantitative | Response Variable Y      - length
# Quantitative Y  | Categorical X (2 Groups)  - length by sex
# Quantitative Y  | Categorical X (3+ Groups) - Nothing in the data with 3+ groups
# Quantitative Y  | Multiple Categorical      - length by sex and biggerfoot
# Quantitative Y  | Quantitative X            - length and width
# Quantitative Y  | Multiple X                - length and width and sex
# Binomial Y      | Quantitative X            - sex and length - opposite x and y 
# Binomial Y      | Multiple X                - Don't worry about it
# Categorical Y   | Categorical X             - biggerfoot and domhand

# C) - table - categorical
table(KidsFeet$biggerfoot, KidsFeet$domhand)

# group_by() - categorical
# summarise() - quantitative

KidsFeet %>% 
  group_by(sex) %>% 
  summarise(mean_length = mean(length))

# D) Histogram   - quantitative
# Box plot       - quantitative
# Dot plot       - quantitative
# Scatter plot   - quantitative
# Bar plot       - categorical

# E) If there is time, try some of the plots in R. Identify the (1) type of data, (2) R Commands, and (3) Graphic that could be used to answer each of the following questions:
# How many boys (B) and how many girls (G) are in the KidsFeet dataset?
# (1) type of data - sex (categorical)
# (2 & 3) table and barplot

table(KidsFeet$sex)

barplot(table(KidsFeet$sex))

# What is the average length of feet in the KidsFeet dataset?
# (1) type of data - length (quantitative)
# (2 & 3) favstats and boxplots

favstats(KidsFeet$length) %>% 
  pander()

boxplot(KidsFeet$length, 
        main="Length of Kids Feet")

ggplot(data=KidsFeet, aes(x=length)) +
  geom_boxplot(fill="navy")+
  labs(
    x="Length of Kids Feet"
  )

# Do boys (B) or girls (G) have longer feet, on average, in the KidsFeet dataset?
# (1) type of data - length (quantitative) and sex (categorical)
# (2 & 3) favstats and boxplots
favstats(length~sex,data=KidsFeet) %>% 
  pander()

boxplot(
        length~sex,
        data=KidsFeet, 
        main="Comparing Length of Feet Across Gender", 
        xlab="Gender",
        ylab="Length of Foot",
        col=c("blue","pink")
        )

ggplot(KidsFeet, aes(x=sex, y=length,fill=sex)) +
  geom_boxplot()+
  labs(title = "Comparing Length of Feet Across Gender",
       x="Gender",
       y="Length of Foot")+
  scale_fill_manual(values=c("blue","pink"))

# Are there certain months of the year that are associated with children having longer feet, on average, than others?
# (1) type of data - length (quantitative) and month (categorical)
# (2 & 3) favstats and boxplots
favstats(length~birthmonth,data=KidsFeet) %>% 
  pander()

boxplot(length~birthmonth,data=KidsFeet, 
        col="navy",
        xlab = "Numerical Birth Month",
        ylab="Feet Length")

ggplot(data=KidsFeet, aes(x=as.factor(birthmonth), y=length)) +
  geom_boxplot(fill="navy")+
  labs(
    x="Numerical Birth Month",
    y="Foot Length"
  )

# Is there a relationship between the length of children's feet and the
# width of their foot?
# (1) type of data - length and width (quantitative)
# (2 & 3) correlation and plot

cor(KidsFeet$length,KidsFeet$width)

plot(length~width, 
      data=KidsFeet, 
      main = "Relationship Between Length and Width of Kids Feet", 
      xlab="Width of Foot", 
      ylab = "Length of Foot")

ggplot(KidsFeet,aes(x=width, y=length))+
  geom_point(color="navy")+
  labs(
    title="Relationship Between Length and Width of Kids Feet",
    x="Width of Foot",
    y="Length of Foot")

ggplot(KidsFeet,aes(x=width,y=length,color=sex))+
  geom_point()+
  labs(title="Relationship Between Length and Width of Kids Feet", 
       x="Width of Foot", 
       y="Length of Foot")+
  scale_color_manual(values=c("blue","pink"))

# Does the width of a child's foot predict whether or not they are right-handed?
# Logistic Regression - will cover later

# Is one gender more likely to be born in certain seasons of the year?
# (1) type of data - season and sex (categorical)
# (2 & 3) table and bar plot

Kids3 <- KidsFeet %>% 
  mutate(
    season = case_when(
      birthmonth %in% c(12,1,2) ~ "Winter",
      birthmonth %in% c(3,4,5) ~ "Spring",
      birthmonth %in% c(6,7,8) ~ "Summer",
      birthmonth %in% c(9,10,11) ~ "Fall"
    )
  )  

pander(table(Kids3$season,KidsFeet$sex))

barplot(table(Kids3$sex, Kids3$season),
        beside = TRUE,
        col = c("blue","pink"),
        legend.text=TRUE, 
        xlab="Handedness", 
        ylab="Frequency", 
        main="Count of Gender by Season")

ggplot(Kids3, aes(x=season,fill=sex))+
  geom_bar(position="dodge")+
  scale_fill_manual(values=c("blue","pink"))


#Make some notes in your index file and knit.
```

## Week 4 Day 1

### T-Test

```{r}
# Libraries used: MOSAIC
# Dataset:        KIDSFEET


# Go to the Making inference page - this first section is referring to book
# Define inference

# Hypothesis Testing
# a.	Type I error vs Type II error – Give an example (Flip flops vs medicine)
# b.	Type I error = Level of significance = 1 – level of confidence
# c.	Type II errors  and Power
# d.	Sufficient evidence – not proof
# e.	P-value need two things: a TEST STATISTIC and a SAMPLING DISTRIBUTION OF TEST STATISTIC

# Find the four parametric distributions (types of tests that are run on each test from 221)

# Parametric methods compared to non-parametric methods
# Non parametric test are still powerful and are likely more powerful than parametric tests if the requirements for a parametric test are not met. 

# An example of doing a t.test
t.test(KidsFeet$length, 
       mu=25.1, 
       alternative="two.sided",
       conf.level=0.95)

# Ho: mu=25.1 Ha:mu ne 25.1
# Level of significance = 0.05
# Test Statistic t=-1.7865 
# df=38 
# p=value = 0.082
# p-value > alpha, therefore we fail to reject the Ho
# We have insufficient evidence to say that the mean is different than 25.1.

pander(t.test(KidsFeet$length, mu=25.1, alternative="less",conf.level=0.95))
# Ho: mu=25.1 Ha:mu < 25.1
# Level of significance = 0.05
# Test Statistic t=-1.7865 df=38 p=value = 0.041
# p-value < alpha, therefore we reject the Ho
# We have sufficient evidence to say that the mean is less than 25.1.

# Note: Because of the Central Limit Theorem, we can assume the distribution of sample means is normal because we have a large sample size (n>=30).

t.test(length~sex,
       data=KidsFeet, 
       mu=0, 
       alternative = "greater",
       conf.level=0.95)

```

## Week 4 Day 2

### T-Test

```{r}
# Libraries used: MOSAIC, PANDER, TIDYVERSE, CAR
# Dataset:        KIDSFEET

# 2)	Have students go through an explanation of what they see in Overview, R instructions, and Explanations.

# ONE SAMPLE T-TEST
pander(t.test(KidsFeet$length, mu=28, alternative="two.sided",conf.level=0.95))

# Ho: mu = 28 Ha: mu ne 28
# $$ H_o: \mu=28 $$
# $$ H_a:\mu \neq 28 $$
# Level of Signficance = 0.05 ($$ \alpha $$)
# Test Statistic t=-15.532 
# DF=38
# p-value < 2.2e-16
# p-value < alpha, therefore we reject the Ho
# We have sufficient evidence to say that the mean is different than 28.

# Note: Because of the Central Limit Theorem, we can assume the distribution of sample means is normal because we have a large sample size (n>30).

# PAIRED T-TEST
KidsFeet3 <- KidsFeet %>% 
  mutate(width3 = 3*width,difference = length -width3) 

t.test(KidsFeet3$length, KidsFeet3$width3, mu=0, paired=TRUE,
       alternative="two.sided",conf.level=0.95)

t.test(KidsFeet3$difference, mu=0, 
       alternative="two.sided",conf.level=0.95)

# Ho: mu_d = 0 Ha: mu_d ne 0
# $$ H_o: \mu_d = 0 H_a:\mu_d \neq 0 $$
# Level of Significance = 0.05 ($$ \alpha $$)
# t = -11.529, df = 38, p-value = 5.668e-14
# p-value < alpha, therefore we reject the Ho
# We have sufficient evidence to say that there is a difference between (width*3) and length.
  
# Note: Because of the Central Limit Theorem, we can assume the distribution of sample means is normal because we have a large sample size (n>30).
  
# INDEPENDENT SAMPLES T-TEST
t.test(length~sex, 
       data=KidsFeet, 
       mu=0,  
       alternative="two.sided",
       conf.level=0.95)

# Ho: mu_boy = mu_girl Ha: mu_boy = mu_girl
# $$ H_o: \mu_boy = \mu_girl H_a:\mu_boy \neq \mu_girl $$
# level of signficance = 0.05 ($$ \alpha $$)
# t = 1.9174, df = 36.275, p-value = 0.06308
# p-value > alpha, therefore we fail to reject the Ho
# We have *insufficient evidence to say that the length is different between boys and girls.
  
# Note: Because of the Central Limit Theorem, we can assume the distribution of sample means is normal because we have a large sample size (n>30).

# CHECKING REQUIREMENTS:
#One Sample t test
qqPlot(KidsFeet$length)
  
#Paired t test
qqPlot(KidsFeet3$length-KidsFeet3$width3)
  
#Independent Samples t test
qqPlot(length~sex,data=KidsFeet)

# GRAPHS:
# One Sample t test
boxplot(KidsFeet$length)
ggplot(data=KidsFeet, aes(x=length))+
  geom_boxplot(fill="blue")

# Paired t test
boxplot(KidsFeet3$length-KidsFeet3$width3)
ggplot(data=KidsFeet3, aes(x=difference))+
  geom_boxplot(fill="blue")

#Independent Samples t test
boxplot(length~sex,data=KidsFeet)
ggplot(data=KidsFeet, aes(y=length, x=sex))+
  geom_boxplot(fill=c("blue","pink"))

# NUMERICAL SUMMARIES
# One Sample t test
pander(favstats(KidsFeet$length))

# Paired t test
pander(favstats(KidsFeet3$length-KidsFeet3$width3))

#Independent Samples t test
pander(favstats(length~sex,data=KidsFeet))

#Put some notes within the the index rmd file
```

## Week 5 Day 1

### Wilcoxon Test

```{r}
# Libraries used: MOSAIC, CAR
# Dataset:        KIDSFEET

# Five steps of doing the Wilcoxon SIGNED-RANK test

# Performing T-TEST
t.test(KidsFeet$length, mu=25)
qqPlot(KidsFeet$length)

# Performing WILCOXON TEST
wilcox.test(KidsFeet$length, 
            mu=25)

# Ho: median = 25
# Ha: median neq 25
# What does the alternative hypothesis means?
# Alternative hypothesis: true location shift is not equal to 0

# Five steps of doing the Wilcoxon RANKED-SUM test (Mann-Whitney)

t.test(length ~ sex, 
       data=KidsFeet)

wilcox.test(length ~ sex, 
            data=KidsFeet)

# The test is negatively affected WHEN THERE ARE TIES (repeated values) present in the data, but the results are STILL USEFUL IF there are relatively few ties.
```

## Week 5 Day 2

### Wilcoxon Test

```{r}
# Question 1
# 1) Function to to use to add a new column - Mutate
# 2) Plots to look at two or more groups    - boxplots and dotplots
# 3) p-value                                - needs test statistic and probability distribution
# 4) Independent Samples t vs Wilcoxon      - Parametric requirements not met? Use Wilcoxon
# 5) Hypothesis
#    Ho: mu1 = mu2  Ha: mu1 ne mu2          - indepdendent samples
#    Ho: difference in medians = 0 Ha:difference in medians ne 0

# Question 2
# 1) Wilcoxon Applet and follow instructions in quiz
# 2) Go through the four questions

# Question 3
# 1) Prepping for next week's assignment
#   https://www.brainhq.com/brain-connection/word-list-recall/
```

## Week 6 Day 1

### One Way ANOVA

```{r}
# Libraries used: MOSAIC, PANDER, TIDYVERSE, LATTICE
# Dataset:        AIRQUALITY

# Find the ANOVA using Statistics Notebook
# Compare the difference between t-test, ANOVA using the index.html file
# Go through the Overview, Explanation, and Instructions in the Textbook

# Using the Airquality data - Define Factor and Level of a Factor
# State briefly that we will cover two-way ANOVA next week

# Use the Airquality Example
# Numerical and Graphical Summary
boxplot(Wind~Month,data=airquality)

ggplot(data=airquality, aes(x=as.factor(Month), y=Wind))+
  geom_boxplot(fill=c("blue","red","yellow","green","purple"))

pander(favstats(Wind~Month,data=airquality))

# Analysis - One Way ANOVA
airq.aov <- aov(Wind~as.factor(Month),data=airquality)

pander(summary(airq.aov))

# Checking Requirements
par(mfrow=c(1,2))

plot(airq.aov,which=1:2)

par(mfrow=c(1,1))

# Use the Chickwt Example
# Null and alternative hypotheses
# $$ H_0:\mu_1 = \mu_2 = \mu_3 = \mu_4 = \mu_5 = \mu_6 $$
# $$ H_a: \text{at least one } \mu \text{ is different} $$
  
# Numerical and Graphical Summary
boxplot(weight~feed,data=chickwts)

ggplot(data=chickwts, aes(y=weight, x=feed))+
  geom_boxplot(fill = c("blue","white","green","yellow","orange","brown"))

xyplot(weight ~ feed, 
      data=chickwts,
      main="Experiment on how Feed Type impacts Chicken Growth",
      ylab="Adult Weight of Chickens (in grams)",
      xlab="Type of Feed Chickens were Fed \n 
      (Blue line shows average weights per feed type.)", 
      type=c("p","a"))

pander(favstats(weight~feed,data=chickwts)[,c("feed","mean","sd","n")])

# Analysis - One Way ANOVA
chickwt.aov <- aov(weight~feed,
                   data=chickwts)

pander(summary(chickwt.aov))

# Checking Requirements
par(mfrow=c(1,2))

plot(chickwt.aov,which=1:2)

# warpbreaks dataset. 

# Question - Are each of the two factors significant and do they interaction
# with each other

# Null and alternative hypotheses
# The first set of null and alternative hypotheses are LISTED FOR EACH FACTOR:

# $$ H_0:\mu_1 = \mu_2 = \mu_3 =\mu \text{for tension}$$
# $$ H_a: \text{at least one } \mu \text{ is different for tension} $$

# $$ H_0:\mu_1 = \mu_2 = \mu \text{for wool}$$
# $$ H_a: \text{at least one } \mu \text{ is different for wool} $$

# $$ H_0:\text{ The effect of tension is the same for all types of wool} $$
# $$ H_a:\text{ The effect of tension is not the same for all types of wool}$$

# install.packages("lattice")


# Numerical Summaries
warpbreaks %>% 
  group_by(tension) %>% 
  summarise(ave=mean(breaks),sd=sd(breaks),SampleSize=n()) %>% 
  pander()

warpbreaks %>% 
  group_by(wool) %>% 
  summarise(ave=mean(breaks),sd=sd(breaks),SampleSize=n()) %>% 
  pander()

warpbreaks %>% 
  group_by(wool, tension) %>% 
  summarise(ave=mean(breaks),sd=sd(breaks),SampleSize=n()) %>% 
  pander()


favstats(breaks~wool+tension,data=warpbreaks)

# Graphical Summaries xyplots 3 plots
xyplot(breaks ~ wool, data=warpbreaks, type=c("p","a"),
       main="Type of Wool", col='blue', 
       xlab="Type of Wool", ylab="Number of Warps that Broke")

xyplot(breaks ~ tension, data=warpbreaks, type=c("p","a"),
       main="Type of Tension", col='blue', 
       xlab="Type of Tension", ylab="Number of Warps that Broke")

xyplot(breaks ~ tension, data=warpbreaks, groups=wool, type=c("p","a"), 
       main="Significance of the Interaction",
       auto.key=list(corner=c(1,1)))

# Inferential Statistics
warp.aov <- aov(breaks ~ wool + tension + wool:tension, data=warpbreaks)

summary(warp.aov) %>% pander()

par(mfrow=c(1,2))

plot(warp.aov, which=1:2, pch=16)

```

## Week 7 Day 2

### Kruskal Wallis

```{r}
# Libraries used: MOSAIC, PANDER, CAR
# Dataset:        KIDSFEET

# Similarities and Differences Between Kruskal-Wallis and Wilcoxon Rank Sum Test
# Hypotheses for Kruskal-Wallis
# Go through explanation of Calculating of Test Statistic - Briefly show the six steps in the explanation
# What can you conclude with a p-value from a Kruskal Wallis
# Kruskal-Wallis Test and Wilcoxon Rank Sum test – If you have multiple samples and not just two, use Kruskal-Wallis test. It is also the non-parametric version of ANOVA.

# Hypotheses

#Kruskal Wallis – go through a few examples below and in the end explain what three codes they need for Kruskal Wallis

par(mfrow=c(1,1))

boxplot(correct ~ condition, data=Friendly, col='grey', ylab="# of Correct Answers", main="# of Correct Answers by Method", xlab="Method")

ggplot(data=Friendly, aes(x=condition, y=correct))+
  geom_boxplot(fill=c("red","white","blue"))

favstats(correct ~ condition, data=Friendly)[,-10] %>%pander(caption="# of Correct Answers by Method")

# Performing ANOVA
sharky.aov <- aov(correct ~ condition,
                  data=Friendly)

summary(sharky.aov)

# Checking Requirements
par(mfrow=c(1,2))

plot(sharky.aov,which=1)

qqPlot(sharky.aov$residuals)

# Performing KRUSKAL-WALIS TEST
kruskal.test(correct ~ condition,
             data=Friendly)

# Use the Chickwt Example
# Numerical and Graphical Summary
par(mfrow=c(1,1))

boxplot(weight~feed,data=chickwts)

ggplot(data=chickwts, aes(x=feed, y=weight))+
  geom_boxplot(fill=c("red","white","blue","red","white","blue"))

pander(favstats(weight~feed,data=chickwts))

# Analysis - One Way ANOVA
chickwt.aov <- aov(weight~feed,
                   data=chickwts)

summary(chickwt.aov)

# Checking Requirements
par(mfrow=c(1,2))

plot(chickwt.aov,which=1)

qqPlot(sharky.aov$residuals)

# Analysis - Kruskal Wallis
kruskal.test(weight~feed,
             data=chickwts)

```

## Week 8 Day 1

### Simple Linear?

```{r}
# Libraries used: MOSAIC, PANDER, TIDYVERSE, DT, CAR
# Dataset:        DAVIS, KIDSFEET

# 1) Students are to find some data to do a simple linear regression for Friday; Need to find two quantitative variables to determine a linear relationship
# 2) Predicted final exam score
# 3) Regression Applet
#   a) Move the sliders on the right - What does "a" represent and what does "b" represent
#   a.	Height to Shoe Size
#   b.	Move sliders and a and b to try and fit the data.
#   c.	Best fit and compare to your line
#   d.	Residuals – squared residuals
#4)	Discuss with neighbor and then with group
#   a.	Changing a? – Explaining the slope

#Talk from Elder Gilbert about Parable of the Slope
#   a) What does least squares regression line mean?

# Open your Statistics-Notebook to the Linear Regression page.See Overview of Simple Linear Regression

# R instructions for Simple Linear Regression and finding the y-intercept and slope
# First one section of Explanation tab
# c. Residual and Errors


# Sample Code for Simple Linear Regression

# Get a Data Set
# Assumptions
# Review some of the explanations

rm("Davis")

## $$ Y_i = \beta_0 +  \beta_1x_i + \epsilon $$
## $$ H_0: \beta_1 = 0 $$
## $$ H_a: \beta_1 \neq 0 $$

##Data Set 1
Davis2 <- Davis %>%
  filter(height>140)

# Now create the model
davis.lm <- lm(height ~ repht, 
               data = Davis2)

predict(davis.lm, 
        newdata = data.frame(repht = 160), 
        type = "response")

summary(davis.lm)

plot(height ~ repht, 
     data = Davis2)

abline(davis.lm,
       lty=1,
       lwd=1)

ggplot(Davis2, aes(x = repht, y = height))+
  geom_point()+
  geom_smooth(method = "lm", 
              formula = y~x, 
              se = FALSE)

DavisNA <- Davis2 %>%
  filter(!is.na(repht),
         !is.na(height))

cor(DavisNA$height,DavisNA$repht)

par(mfrow=c(1,3))

plot(davis.lm,which=1)

qqPlot(davis.lm$residuals)

plot(davis.lm$residuals)

##Data Set 2
kidsfeet.lm <- lm(length ~ width, 
                  data = KidsFeet)

summary(kidsfeet.lm)

plot(length ~ width, 
     data = KidsFeet)

abline(kidsfeet.lm,
       lty=1,
       lwd=1)

ggplot(KidsFeet, aes(x = width, 
                     y = length))+
  geom_point()+
  geom_smooth(method = "lm", 
              formula = y~x, 
              se = FALSE)

cor(KidsFeet$length, KidsFeet$width)

par(mfrow=c(1,3))

plot(kidsfeet.lm,which=1:2)

plot(kidsfeet.lm$residuals)

# The Mathematical Model of Regression (including the equation of the best-fit line).
# How to interpret the slope and intercept in a Simple Linear Regression.
# What a residual is in a regression.
# How to measure the correlation of a regression.
# How to diagnose the appropriateness of a regression for a given data set.
# How to perform a regression in RStudio and locate the slope and intercept from the output summary.

```

## Week 8 Day 2

### Simple Linear?

```{r}
# Libraries used: MOSAIC, PANDER, TIDYVERSE, DT, CAR
# Dataset:        KIDSFEET

# Go through the four questions in the class quiz with at least one other
# person and use the simple linear regression section within the textbook
# You can use the code below to practice items with a simple linear regression.

# Kids Feet
# $$ Y_i = \beta_0 +  \beta_1x_i + \epsilon $$
# $$ H_0: \beta_1 = 0 $$
# $$ H_a: \beta_1 \neq 0 $$

kidsfeet.lm <- lm(length ~ width, 
                  data = KidsFeet)

summary(kidsfeet.lm)

plot(length ~ width, 
     data = KidsFeet)

abline(kidsfeet.lm,
       lty=1,
       lwd=1)

ggplot(KidsFeet, aes(x = width, y = length))+
  geom_point()+
  geom_smooth(method = "lm", 
              formula = y~x, 
              se = FALSE)

par(mfrow=c(1,3))

plot(kidsfeet.lm,
     which=1:2)

plot(kidsfeet.lm$residuals)
```

## Week 9 Day 1

### Multiple Linear?

```{r}
# Libraries used: TIDYVERSE, CAR
# Dataset:        MTCARS

# Go through the example using the TWO LINE MODEL.
# 1)	State the mathematical equation of two-line model
# 2) lm.2lines <-lm(mpg ~ qsec + am + qsec:am, data=mtcars) -  
#   Use the example in the textbook to go through what the estimates mean:
# 3)	Interpret the intercept and slope of each of the two lines (the answer is below)
# 4)	Create a scatterplot using the code below

#Full Model
lm.2lines <-lm(mpg ~ qsec + am + qsec:am, 
               data=mtcars)

summary(lm.2lines)

lm.2lines <-lm(mpg ~ qsec*am, 
               data=mtcars)

summary(lm.2lines)

#Equal Slopes Model
eqslopes <-lm(mpg ~ qsec + am, 
              data=mtcars)

summary(eqslopes)

#Equal Intercepts Model
eqint <-lm(mpg ~ qsec + qsec:am, 
           data=mtcars)

summary(eqint)

lm.line <-lm(mpg ~ qsec, 
             data=mtcars)
summary(lm.line)

#Create graph using the following code:
par(mfrow=c(1,1))

plot(mpg ~ qsec, 
    data=mtcars, 
    col=c("yellow","red")[as.factor(am)], 
    pch=16, 
    bg="gray83", 
    main="Two-lines Model using mtcars data set", 
    cex.main=1)
legend("topleft", legend=c("Baseline (am==0)", "Changed-line (am==1)"), 
bty="n", lty=1, col=c("yellow","red"), cex=0.8)

# Get the "Estimates" automatically:
# Then b will have 4 estimates:
# b[1] is the estimate of beta_0: -9.0099
# b[2] is the estimate of beta_1:  1.4385
# b[3] is the estimate of beta_2: -14.5107
# b[4] is the estimate of beta_3: 1.3214

# Full Model  - redo graph first
b <- coef(lm.2lines)
curve(b[1] + b[2]*x, 
      col="yellow", 
      lwd=2, 
      add=TRUE)  #baseline (in blue)
curve((b[1] + b[3]) + (b[2] + b[4]) * x, 
      col="red", 
      lwd=2, 
      add=TRUE) #changed line (in orange)
b

# Equal Slopes Model - redo graph first
b <- coef(eqslopes)
curve(b[1] + b[2]*x, col="yellow", 
      lwd=2, 
      add=TRUE)  #baseline (in blue)
curve((b[1] + b[3]) + (b[2]) * x, 
      col="red", 
      lwd=2, 
      add=TRUE) #changed line (in orange)

# Equal Intercepts Model  - redo graph first
b <- coef(eqint)
curve(b[1] + b[2]*x, 
      col="yellow", 
      lwd=2, 
      add=TRUE)  #baseline (in blue)

curve((b[1]) + (b[2] + b[3]) * x, 
      col="red", 
      lwd=2, 
      add=TRUE) #changed line (in orange)

# Then b will have 4 estimates:
# b[1] is the estimate of beta_0: -9.0099
# b[2] is the estimate of beta_1:  1.4385
# b[3] is the estimate of beta_2: -14.5107
# b[4] is the estimate of beta_3: 1.3214

# Full Model Graph
b <- coef(lm.2lines)
ggplot(mtcars, aes(y=mpg, x=qsec, color=factor(am))) +
  geom_point(pch=16) +
  #geom_smooth(method="lm", se=F) + #easy way, but only draws the full interaction model. The manual way using stat_function (see below) is more involved, but more dynamic.
  stat_function(fun = function(x) b[1] + b[2]*x, color="red") + #am==0 line
  stat_function(fun = function(x) (b[1]+b[3]) + (b[2]+b[4])*x,color="yellow") + #am==1 line 
  scale_color_manual(name="Transmission (am)", values=c("red","yellow")) +
  labs(title="Two-lines Model using mtcars data set")

# Equal Slopes Graph
b <- coef(eqslopes)

ggplot(mtcars, aes(y=mpg, x=qsec, color=factor(am))) +
  geom_point(pch=16) +
  # geom_smooth(method="lm", se=F) + #easy way, but only draws the full interaction model. The manual way using stat_function (see below) is more involved, but more dynamic.
  stat_function(fun = function(x) b[1] + b[2]*x, color="red") + #am==0 line
  stat_function(fun = function(x) (b[1]+b[3]) + (b[2])*x,color="yellow") + #am==1 line 
  scale_color_manual(name="Transmission (am)", values=c("red","yellow")) +
  labs(title="Two-lines Model using mtcars data set")

# Equal Intercepts Graph
b <- coef(eqint)
ggplot(mtcars, aes(y=mpg, x=qsec, color=factor(am))) +
  geom_point(pch=16) +
  #geom_smooth(method="lm", se=F) + #easy way, but only draws the full interaction model. The manual way using stat_function (see below) is more involved, but more dynamic.
  stat_function(fun = function(x) b[1] + b[2]*x, color="red") + #am==0 line
  stat_function(fun = function(x) (b[1]) + (b[2]+b[3])*x,color="yellow") + #am==1 line 
  scale_color_manual(name="Transmission (am)", values=c("red","yellow")) +
  labs(title="Two-lines Model using mtcars data set")


ggplot(mtcars, aes(y=mpg, x=qsec, color=factor(am))) +
  geom_point(pch=16, bg="gray83") +
  geom_smooth(method="lm", se=F) + #easy way, but only draws the full interaction model. The manual way using stat_function (see below) is more involved, but more dynamic.
  scale_color_manual(name="Transmission (am)", values=c("red","yellow")) +
  labs(title="Two-lines Model using mtcars data set")


# Check the requirements to see if the model is good
par(mfrow=c(1,3))
plot(lm.2lines,which=1)

qqPlot(lm.2lines$residuals)
plot(lm.2lines$residuals)

# Check for significance
# qsec is significant  and the change of slope is significant

```

## Week 9 Day 2

### Multiple Linear?

```{r}
# 1)	Each student draw a sketch of the two-line model and explain to one other person
# 2) LaTex code for the model and null and alternative hypotheses
# $$   Y_i = \beta_0 + \beta_1{X_{i1}} + \beta_2{X_{i2}} + \beta_3{x_{i1}}{x_{i2}} + \epsilon_i $$
# $$ H_0: \beta_1 = 0 $$
# $$ H_a: \beta_1 \neq 0 $$
# $$ H_0: \beta_2 = 0 $$
# $$ H_a: \beta_2 \neq 0 $$
# $$ H_0: \beta_3 = 0 $$
# $$ H_a: \beta_3 \neq 0 $$
# 2) Go through one example of the three and then have them try the other two:

#Equal slopes
par(mfrow=c(1,1))

palette(c("skyblue","firebrick"))

plot(mpg ~ qsec, data=mtcars, col=as.factor(am), xlim=c(0,30), ylim=c(-30,40), main="1974 Motor Trend Cars", pch=16)

equalslopes.lm <- lm(mpg~qsec + as.factor(am), data=mtcars)

summary(equalslopes.lm)

b <- coef(equalslopes.lm)

b

abline( b[1], b[2], col=palette()[1])
abline( b[1]+b[3],b[2], col=palette()[2])

legend("topleft", legend=c("automatic","manual"), pch=1, col=palette(), title="Transmission (am)", bty="n")
# Note:  You may need to increase the graph window so the legend is in the right place.

# Equal intercepts

palette(c("skyblue","firebrick"))

plot(mpg ~ qsec, data=mtcars, col=as.factor(am), xlim=c(0,30), ylim=c(-30,40), main="1974 Motor Trend Cars", pch=16)

equalint.lm <- lm(mpg~qsec + qsec:as.factor(am), data=mtcars)

summary(equalint.lm)

b <- coef(equalint.lm)

abline( b[1], b[2], col=palette()[1])
abline( b[1],b[2]+b[3], col=palette()[2])
legend("topleft", legend=c("automatic","manual"), pch=1, col=palette(), title="Transmission (am)", bty="n")

#Full Model
palette(c("skyblue","firebrick"))
plot(mpg ~ qsec, data=mtcars, col=as.factor(am), xlim=c(0,30), ylim=c(-30,40), main="1974 Motor Trend Cars", pch=16)

fullmodel.lm <- lm(mpg~qsec + as.factor(am) + as.factor(am):qsec, data=mtcars)
summary(fullmodel.lm)

b <- coef(fullmodel.lm)

abline( b[1], b[2], col=palette()[1])
abline( b[1]+b[3],b[2]+b[4], col=palette()[2])

legend("topleft", legend=c("automatic","manual"), pch=1, col=palette(), title="Transmission (am)", bty="n")

# Summarize your Findings
# What have you learned? Discuss the following with your peers as you complete this activity.
# Are all p-values significant in your lm(...) for the full model?
# How about for the equal intercepts model?
# How about for the equal slopes model?
```

## Week 10 Day 1

### Logistic Regression?

```{r}
# Libraries used: MOSAIC, PANDER, TIDYVERSE, DT, ALR4, RESOURCESELECTION
# Dataset:        CHALLENGER

# Open Challenger example and see if you can watch a few minutes from the video
# Look at the second paragraph on how the data was collected, then go to the explanation of the logistic model and explain:
# a.The model
# b.The log of the odds

chall.glm <- glm(fail>0 ~ temp, 
                 data=Challeng, 
                 family=binomial)

summary(chall.glm) %>% 
  pander()

# Plotting the model
plot( fail>0 ~ temp, data=Challeng, 
      xlab = "Outside Temperature at Time of Launch (Fahrenheit)",
      col  = c("green","red")[as.factor(fail>0)],
      ylab ='Probability of At least One O-ring Failing', pch=16, 
      main = "NASA Shuttle Launch Data from 1981 to 1985", xlim=c(30,85))

# Getting the coeffiencts
b <- coef(chall.glm)
b

curve(exp(b[1]+b[2]*x)/(1+exp(b[1]+b[2]*x)), add=TRUE)

# Testing the fit of the model
hoslem.test(chall.glm$y, chall.glm$fitted, g=6) %>% 
  pander()

# If multiple repeated x values then use the following:
pchisq(20.32, 21, lower.tail=FALSE)

# Prediciting a failure based on a temperature: 
predict(chall.glm, newdata = data.frame(temp = 64.75), type = "response")
#0.9996088

# Interpreting exp(b1) shows that the odds of the o-rings failing for a given launch decreases by a factor of 0.79 for every 1° F increase in temperature. Said differently, the odds of an o-ring failure during launch decrease by 21% (1-0.79) for every 1° F increase in temperature. (Also, from the reverse perspective, every 1° F decrease in temperature increases the odds of a failed o-ring by a factor of exp(-0.2322) exp(b[2])
#0.79

#Add some notes to your Table of Contents about logistic regression.

# Things to do Further:
# 1) Probability on y-axis
# 2) Binomial is using logit equation
# 3) Checking requirements
# 4) Explain of each of the items in the equation mean

# Another example that I cannot run:
# crime.lm <- lm(crimerate ~ exams, data=andrew)
# summary(crime.lm)
# crime.glm <- glm((crimerate > 210.5)~exames, data=andrew, family=binomial)
# summary(crime.glm)

```

## Week 10 Day 2

### Logistic Regression?

```{r}
# Libraries used: PANDER, TIDYVERSE, RESOURCESELECTION
# Dataset:        KIDSFEET

# Do the model: 
kidsfeet.glm <- glm(sex=="B" ~ length, 
                    data=KidsFeet, 
                    family=binomial)

summary(kidsfeet.glm)

#Plotting the model
plot( sex=="B" ~ length, 
      data=KidsFeet, 
      xlab="Length",
      col = c("red","lightblue")[as.factor(sex=="B")],
      ylab="Probability of of a Child being a Boy", 
      pch=16, 
      main="KidsFeet Logistic Regression")

b <- coef(kidsfeet.glm)
b

curve(exp(b[1]+b[2]*x)/(1+exp(b[1]+b[2]*x)), add=TRUE)

hoslem.test(kidsfeet.glm$y, kidsfeet.glm$fitted, g=6) %>% 
  pander()

# Predict 
predict(kidsfeet.glm, 
        newdata = data.frame(length = 25), 
        type = "response")

# Calculating odds
exp(b[2])
#1.660921

# Interpreting
# If beta1 is significant then the explanatory variable does well in predicting the odds of a “success”.  But interpreting e(b1) is more important.  As the length of a foot in cm goes up by one, the odds of the child being a boy increases by a factor of 0.66 or 66%. 

```

## Week 12 Day 1

### Chi-squared Test

```{r}
# Libraries used: TIDYVERSE
# Dataset:        MTCARS, FRUIT

# Go through the combine function in the textbook
Fruitrow <- rbind( 
                  Apples = c(Good = 80, 
                              Bruised = 20, 
                              Rotten = 15), 
                  Oranges = c(Good = 75, 
                               Bruised = 25, 
                               Rotten = 10)
                  )
Fruitrow
#          Good Bruised Rotten
# Apples    80      20     15
# Oranges   75      25     10

Fruitcolumn <- cbind(
                    Apples = c(Good = 80, 
                               Bruised = 20, 
                               Rotten = 15), 
                    Oranges = c(Good = 75, 
                                Bruised = 25, 
                                Rotten = 10) )

Fruitcolumn

#         Apples Oranges
#Good        80      75
#Bruised     20      25
#Rotten      15      10

# 5)	Go through an example of a barplot
# 6)	Note: Rows are legends, and columns are columns or bars

barplot(Fruitrow,beside = TRUE, legend.text=TRUE)
barplot(Fruitcolumn,beside = TRUE, legend.text=TRUE)

# 9) Using data
table(mtcars$am, mtcars$cyl)

ggplot(data=mtcars,aes(x=as.factor(cyl),
                       fill=as.factor(am)))+
  geom_bar(stat = "count", position = position_dodge())+
  scale_fill_manual(values = c("red","green"))+
  labs(
    x="cylinder",
    y="Number of Cars",
    fill="Transmission"
  )

# 10) Using data
barplot(table(mtcars$am, mtcars$cyl), 
        beside=TRUE, 
        col=c("red","green"), 
        legend.text=TRUE, 
        xlab="Cylinders")

# Doing chisq analysis – chisq.test(x) using both data
Fruitchi <- chisq.test(Fruitcolumn)
Fruitchi

mtchi <- chisq.test(table(mtcars$am, mtcars$cyl))
mtchi

# 12)	Check the requirements using mychisq$expected and mean(mychisq$expected) for both data
Fruitchi$expected
#             Good Bruised   Rotten
#Apples  79.22222      23 12.77778
#Oranges 75.77778      22 12.22222 

mtchi$expected
#       4       6      8
#0 6.53125 4.15625 8.3125
#1 4.46875 2.84375 5.6875

mean(mtchi$expected)
#[1] 5.333333

# 13)	Do Pearsons residuals on both data – mychisq$residuals
Fruitchi$residuals
mtchi$residuals

```

## Week 12 Day 2

### Chi-squared Test

```{r}
# Libraries used: TIDYVERSE
# Dataset:        MTCARS, FRUIT

# 1)	Open statistics notebook and go through the combine function in the textbook
Fruitrow <- rbind( Apples = c(Good = 80, 
                              Bruised = 20, 
                              Rotten = 15), 
                   Oranges = c(Good = 75, 
                               Bruised = 25, 
                               Rotten = 10) )
Fruitrow

#          Good Bruised Rotten
#Apples    80      20     15
#Oranges   75      25     10

Fruitcolumn <- cbind( Apples = c(Good = 80, 
                                 Bruised = 20, 
                                 Rotten = 15), 
                      Oranges = c(Good = 75, 
                                  Bruised = 25, 
                                  Rotten = 10) )
Fruitcolumn

#         Apples Oranges
# Good        80      75
# Bruised     20      25
# Rotten      15      10

# 5)	Go through an example of a barplot
# 6)	Note: Rows are legends, and columns are columns or bars
barplot(Fruitrow,beside = TRUE, legend.text=TRUE)
barplot(Fruitcolumn,beside = TRUE, legend.text=TRUE)

# 9) Using data
table(mtcars$am, mtcars$cyl)

ggplot(data=mtcars,aes(x=as.factor(cyl),fill=as.factor(am)))+
  geom_bar(stat = "count",position = position_dodge())+
  scale_fill_manual(values = c("red","green"))+
  labs(
    x="cylinder",
    fill="Transmission",
    y="Number of Cars"
  )

# 10) Using data
barplot(table(mtcars$am, mtcars$cyl), beside=TRUE, 
        col=c("red","green"), legend.text=TRUE, xlab="Cylinders")

# Doing chisq analysis – chisq.test(x) using both data
Fruitchi <- chisq.test(Fruitcolumn)
Fruitchi
mtchi <- chisq.test(table(mtcars$am, mtcars$cyl))
mtchi

# 12)	Check the requirements using mychisq$expected and mean(mychisq$expected) for both data
Fruitchi$expected

#             Good Bruised   Rotten
#Apples  79.22222      23 12.77778
#Oranges 75.77778      22 12.22222 

mtchi$expected
#       4       6      8
#0 6.53125 4.15625 8.3125
#1 4.46875 2.84375 5.6875

mean(mtchi$expected)
#[1] 5.333333
  
#13)	Do Pearsons residuals on both data – mychisq$residuals
Fruitchi$residuals
mtchi$residuals
```
